### 2018-12-08 paper.tex
Add latex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{multicol}
\setlength{\columnsep}{1cm}

\begin{document}
    \begin{center}
        \textbf{Evaluation of Classifiers on Spam Filtering} 
    \end{center}

    \begin{center}
        Griffin Dunn and William Thompson
    \end{center}

    \textbf{Abstract:} Spam Filters rely on a variety of techniques to work effectively.

    \begin{multicols}{2}
        \section{Introduction}
    \end{multicols}

\end{document}